---
title: "Step 1: Create food license shapefile"
author: "Laura Hatt"
date: '2023-05-03'
output: html_document
---

Inspired by: https://voorheescenter.uic.edu/news-stories/what-was-covids-impact-on-chicagos-food-industry/ 

*Step 1: Load packages and data*

```{r}
library(tidyverse)
library(sf)
#library(tmap)
#library(tmaptools)
library(RSocrata)
library(tidygeocoder)
#library(terra)
#library(raster)

path = "/Users/laurahatt/Documents/GitHub/chicago_food_licenses/"
```

Community Area Boundaries
```{r}
#download from: https://data.cityofchicago.org/Facilities-Geographic-Boundaries/Boundaries-Community-Areas-current-/cauq-8yn6
comm_areas <- read_sf("https://data.cityofchicago.org/api/geospatial/cauq-8yn6?method=export&format=GeoJSON")

#remove unnecessary columns
comm_areas <- comm_areas %>% dplyr::select("community")
```


Business Licenses - Current Active
```{r}
#download from: https://data.cityofchicago.org/Community-Economic-Development/Business-Licenses-Current-Active/uupf-x98q
licenses <- read_csv(paste0(path, "business_licenses.csv"))

#remove spaces from column names
names(licenses) <- make.names(names(licenses),unique = TRUE)
```

*Step 2: Filter to include only Chicago food-related business licenses*

```{r}
#filter to include only business licenses in Chicago
licenses <- filter(licenses, CITY == "CHICAGO")
```



The next part involves some discretion.

Per the Voorhees report:
The food business classifications are as follows:
• Catering
• Shared Kitchens
• Packaged Goods
• Food Peddlers
• Retail Food
• Special Event Food & Liquor
• Tavern
• Wholesale Food Establishment

However, when I review the unique strings in the "License Description" column, I see the following food-related categories:

```{r}
food_related_license_types <- c("Retail Food Establishment", 
                                "Mobile Food License", 
                                "Wholesale Food Establishment", 
                                "Special Event Food",
                                "Shared Kitchen User (Long Term)",
                                "Shared Kitchen User (Short Term)",
                                "Food - Shared Kitchen",
                                "Food - Shared Kitchen - Supplemental",
                                "Pop-Up Establishment Host - Tier II",
                                "Pop-Up Establishment Host - Tier III",
                                "Pop-Up Food Est. User - Tier I",
                                "Pop-Up Food Est. User - Tier II",
                                "Pop-Up Food Est. User - Tier III",
                                "Outdoor Patio",
                                "Package Goods",
                                "Tavern",
                                "Caterer's Liquor License" )

food_licenses <- filter(licenses, LICENSE.DESCRIPTION %in% food_related_license_types)
```

*Step 3: Use coordinates to create sf object*

Geocode licenses that are missing coordinates:
```{r}
#collect rows with missing coordinates
missing_coords <- food_licenses[is.na(food_licenses$LOCATION),]

#geocode addresses for rows with missing coordinates
missing_coords_geocoded <- missing_coords %>%
  geocode_combine(queries = list(list(method = "census", mode = "batch"), 
                                 list(method = "census", mode = "single"),
                                 list(method = "osm")),
                  global_params = list(street = "ADDRESS", 
                                       city = "CITY",
                                       state = "STATE", 
                                       postalcode = "ZIP.CODE"),
                  query_names = c("census_batch", "census_single", "osm"))

#remove rows with "problem addresses" that the geocoder couldn't manage
#inspect later for trends
missing_coords_geocoded <- missing_coords_geocoded %>% drop_na(lat)

#move geocoder lat and long into original columns and delete new columns
missing_coords_geocoded$LATITUDE <- missing_coords_geocoded$lat
missing_coords_geocoded$LONGITUDE <- missing_coords_geocoded$long
missing_coords_geocoded <- subset(missing_coords_geocoded, select = -c(lat, long, query))
```

Add geocoded coordinates to the dataframe:
```{r}
#remove rows with missing coordinates from main dataset
food_licenses <- food_licenses %>% drop_na(LOCATION)

#add rows with geocoded coordinates
food_licenses_coords <- rbind(food_licenses, missing_coords_geocoded)
```

Create SF object:
```{r}
food_licenses_sf <- st_as_sf(food_licenses_coords, coords = c("LONGITUDE","LATITUDE"), crs = 4326) 
```

Eliminate unnecessary columns, for simplicity (for now)
```{r}
food_licenses_sf <- food_licenses_sf %>% select(LICENSE.ID, LEGAL.NAME, ADDRESS)
```

Save as shapefile
```{r}
write_sf(obj=food_licenses_sf, dsn=paste0(path, "Shapefiles/food_licenses.shp"))
```

